### 2021-03-17 DataSelection_EDA.RMD
data sets exploration - slight update on Project instruction

---
title: "MSDS6372_EDAProject2"
author: "Mingyang Nick YU"
date: "3/17/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## This is EDA for MSDS 6372 Project 2

#### First part below is for dataset selection

1. Predicting if a customer will subscribe to a term deposit
Resource:
https://archive.ics.uci.edu/ml/datasets/Bank+Marketing
Pro: No missing data, all missing is marked as unknown - Dataset is fairly clean
Con: No: 36548, Yes:4640 - Heavy unbalanced data 9:1
```{r}
setwd("/Users/mingyang/Desktop/SMU/Applied Statistics/MSDS6372Project2/bank-additional")
library(tidyverse)
bank = read.csv("bank-additional-full.csv",header=TRUE,sep=";") 
str(bank)
variables.to.factor = c("job","marital","education","default","housing","loan","contact","month","day_of_week","poutcome","y")
bank[variables.to.factor] = lapply(bank[variables.to.factor],factor)
summary(bank)
summary(bank$job)
bank %>% filter_all(any_vars(is.na(.)))
```

2. Predicting if someone makes over 50k
https://archive.ics.uci.edu/ml/datasets/Adult
Pro: Easier to understand data, less variable, Train Test set already split and provided for you
Con: Have missing data - however due to the amount we can either delete or set them to unknown (should not affect result as much)
    Still have unbalance data for training dataset <=50k: 24720, >50kL 7841 - not as bad
                                  test dataset <=50K.:12435 >50K. : 3846
```{r}
setwd("/Users/mingyang/Desktop/SMU/Applied Statistics/MSDS6372Project2/Adult-Income")
library(tidyverse)
adult = read.csv("adult.data.csv",header=FALSE,sep=",",strip.white=TRUE) 
colnames(adult) <- c(
  "age", # continuous
  "workclass",# Private, Self-emp-not-inc, Self-emp-inc, Federal-gov, Local-gov, State-gov, Without-pay, Never-worked
  "fnlwgt", # continous, final weight. In other words, this is the number of people the census believes the entry represents
  "education", # Bachelors, Some-college, 11th, HS-grad, Prof-school, Assoc-acdm, Assoc-voc, 9th, 7th-8th, 12th, 
              # Masters, 1st-4th, 10th, Doctorate, 5th-6th, Preschool.education-num: continuous
  "education-num", # continuous
  "marital-status", # Married-civ-spouse, Divorced, Never-married, Separated, Widowed, Married-spouse-absent, Married-AF-spouse.
  "occupation",  # Tech-support, Craft-repair, Other-service, Sales, Exec-managerial, Prof-specialty, Handlers-cleaners, Machine-op-inspct,                 # Adm-clerical, Farming-fishing, Transport-moving, Priv-house-serv, Protective-serv, Armed-Forces
  "relationship", # Wife, Own-child, Husband, Not-in-family, Other-relative, Unmarried
  "race", # White, Asian-Pac-Islander, Amer-Indian-Eskimo, Other, Black
  "sex",   # Female, Male
  "capital-gain", # continuous
  "capital-loss", # continuous
  "hours-per-week", # continuous
  "native-country", # United-States, Cambodia, England, Puerto-Rico, Canada, Germany, Outlying-US(Guam-USVI-etc), India, Japan, Greece, South, China, Cuba, Iran, Honduras, Philippines, Italy, Poland, Jamaica, Vietnam, Mexico, Portugal, Ireland, France, Dominican-Republic, Laos, Ecuador, Taiwan, Haiti, Columbia, Hungary, Guatemala, Nicaragua, Scotland, Thailand, Yugoslavia, El-Salvador, Trinadad&Tobago, Peru, Hong, Holand-Netherlands
  "income" # expects to predict
)
str(adult)
## First, convert "?"s to NAs...
adult[adult == "?"] <- NA
variables.to.factor = c("workclass","education","marital-status","occupation","relationship","race","sex","native-country","income")
adult[variables.to.factor] = lapply(adult[variables.to.factor],factor)
summary(adult)

# inporting testing data
adult.test = read.csv("adult.test.csv",header=FALSE,sep=",",strip.white=TRUE)
colnames(adult.test) <- c(
  "age", # continuous
  "workclass",# Private, Self-emp-not-inc, Self-emp-inc, Federal-gov, Local-gov, State-gov, Without-pay, Never-worked
  "fnlwgt", # continous, final weight. In other words, this is the number of people the census believes the entry represents
  "education", # Bachelors, Some-college, 11th, HS-grad, Prof-school, Assoc-acdm, Assoc-voc, 9th, 7th-8th, 12th, 
              # Masters, 1st-4th, 10th, Doctorate, 5th-6th, Preschool.education-num: continuous
  "education-num", # continuous
  "marital-status", # Married-civ-spouse, Divorced, Never-married, Separated, Widowed, Married-spouse-absent, Married-AF-spouse.
  "occupation",  # Tech-support, Craft-repair, Other-service, Sales, Exec-managerial, Prof-specialty, Handlers-cleaners, Machine-op-inspct,                 # Adm-clerical, Farming-fishing, Transport-moving, Priv-house-serv, Protective-serv, Armed-Forces
  "relationship", # Wife, Own-child, Husband, Not-in-family, Other-relative, Unmarried
  "race", # White, Asian-Pac-Islander, Amer-Indian-Eskimo, Other, Black
  "sex",   # Female, Male
  "capital-gain", # continuous
  "capital-loss", # continuous
  "hours-per-week", # continuous
  "native-country", # United-States, Cambodia, England, Puerto-Rico, Canada, Germany, Outlying-US(Guam-USVI-etc), India, Japan, Greece, South, China, Cuba, Iran, Honduras, Philippines, Italy, Poland, Jamaica, Vietnam, Mexico, Portugal, Ireland, France, Dominican-Republic, Laos, Ecuador, Taiwan, Haiti, Columbia, Hungary, Guatemala, Nicaragua, Scotland, Thailand, Yugoslavia, El-Salvador, Trinadad&Tobago, Peru, Hong, Holand-Netherlands
  "income" # expects to predict
)
## First, convert "?"s to NAs...
adult.test[adult.test == "?"] <- NA
variables.to.factor = c("workclass","education","marital-status","occupation","relationship","race","sex","native-country","income")
adult.test[variables.to.factor] = lapply(adult.test[variables.to.factor],factor)
summary(adult.test)

adult.missing = adult %>% filter_all(any_vars(is.na(.)))
adult.test.missing = adult.test %>% filter_all(any_vars(is.na(.)))
view(adult.missing)
```

3. Assessing risk factors and predicting if a woman with osteopetrosis will have a bone fracture within the first year after joining the study.
Resource:R package aplore3, use the glow_bonemed data set
Pro: No missing data, already cleaned data for the most part, less many level categories
Con: Much smaller dataset - ratio of fracture is still about 3:1 No:Yes
```{r}
library(aplore3)
head(glow_bonemed)
str(glow_bonemed)
summary(glow_bonemed)
```

### 2021-03-20 DataSelection_EDA.RMD
update DataSelection

@@
   "fnlwgt", # continous, final weight. In other words, this is the number of people the census believes the entry represents
-  "education", # Bachelors, Some-college, 11th, HS-grad, Prof-school, Assoc-acdm, Assoc-voc, 9th, 7th-8th, 12th, 
+  "education", # Bachelors, Some-college, //11th, HS-grad, Prof-school, Assoc-acdm, Assoc-voc, 9th, 7th-8th, 12th, 
               # Masters, 1st-4th, 10th, Doctorate, 5th-6th, Preschool.education-num: continuous
@@
 adult.test.missing = adult.test %>% filter_all(any_vars(is.na(.)))
-view(adult.missing)
+# Look at missing data
+library(naniar)
+gg_miss_var(adult)
+gg_miss_var(adult.test)
 ```
@@
 summary(glow_bonemed)
+gg_miss_var(glow_bonemed)
 ```

### 2021-03-23 MSDS6372Project2.Rmd
add initial EDA and image files

---
title: "MSDS6372Project2"
author: "Amber Clark, Feby Thomas Cheruvathoor and Mingyang Nick YU"
date: "3/22/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Data Import and Initial Inspection
```{r}
setwd("/Users/mingyang/Desktop/SMU/Applied Statistics/MSDS6372Project2/Adult-Income")
library(tidyverse)
adult = read.csv("adult.data.csv",header=FALSE,sep=",",strip.white=TRUE) 
# Assign column names
colnames(adult) <- c(
  "age", # continuous
  "workclass",# Private, Self-emp-not-inc, Self-emp-inc, Federal-gov, Local-gov, State-gov, Without-pay, Never-worked
  "fnlwgt", # continous, final weight. In other words, this is the number of people the census believes the entry represents
  "education", # Bachelors, Some-college, 11th, HS-grad, Prof-school, Assoc-acdm, Assoc-voc, 9th, 7th-8th, 12th, 
              # Masters, 1st-4th, 10th, Doctorate, 5th-6th, Preschool.education-num: continuous
  "education.num", # continuous
  "marital.status", # Married-civ-spouse, Divorced, Never-married, Separated, Widowed, Married-spouse-absent, Married-AF-spouse.
  "occupation",  # Tech-support, Craft-repair, Other-service, Sales, Exec-managerial, Prof-specialty, Handlers-cleaners, Machine-op-inspct,                 # Adm-clerical, Farming-fishing, Transport-moving, Priv-house-serv, Protective-serv, Armed-Forces
  "relationship", # Wife, Own-child, Husband, Not-in-family, Other-relative, Unmarried
  "race", # White, Asian-Pac-Islander, Amer-Indian-Eskimo, Other, Black
  "sex",   # Female, Male
  "capital.gain", # continuous
  "capital.loss", # continuous
  "hours.per.week", # continuous
  "native.country", # United-States, Cambodia, England, Puerto-Rico, Canada, Germany, Outlying-US(Guam-USVI-etc), India, Japan, Greece, South, China, Cuba, Iran, Honduras, Philippines, Italy, Poland, Jamaica, Vietnam, Mexico, Portugal, Ireland, France, Dominican-Republic, Laos, Ecuador, Taiwan, Haiti, Columbia, Hungary, Guatemala, Nicaragua, Scotland, Thailand, Yugoslavia, El-Salvador, Trinadad&Tobago, Peru, Hong, Holand-Netherlands
  "income" # expects to predict
)
str(adult)
## First, convert "?"s to Not-Known...
adult[adult == "?"] <- "Not-Known"

# Decision is made to delete Not-Known observations under workclass & occupation (based on EDA below)
# 1. workclass and occupation missing values are almost identical - except a few workclass assigned Never-worked has a few Not-Known
# workclass and occupation seem to both being good potential predictors for response income
# Thus decision is made to delete Not-Known under occupation and workclass - restrict our range to only predict observation that has occupation or workclass that is known to us
adult = adult%>% filter(workclass!="Not-Known" & occupation!="Not-Known")

# Reorder education levels for later
edu.levels = factor(c("Preschool","1st-4th","5th-6th","7th-8th","9th","10th","11th","12th","HS-grad","Some-college","Assoc-voc","Assoc-acdm","Bachelors","Masters","Prof-school","Doctorate"))
adult$education <- factor(adult$education, levels = edu.levels)
variables.to.factor = c("workclass","marital.status","occupation","relationship","race","sex","native.country","income")
adult[variables.to.factor] = lapply(adult[variables.to.factor],factor)
summary(adult)

# inporting testing data
adult.test = read.csv("adult.test.csv",header=FALSE,sep=",",strip.white=TRUE)
colnames(adult.test) <- c(
  "age", # continuous
  "workclass",# Private, Self-emp-not-inc, Self-emp-inc, Federal-gov, Local-gov, State-gov, Without-pay, Never-worked
  "fnlwgt", # continous, final weight. In other words, this is the number of people the census believes the entry represents
  "education", # Bachelors, Some-college, 11th, HS-grad, Prof-school, Assoc-acdm, Assoc-voc, 9th, 7th-8th, 12th, 
              # Masters, 1st-4th, 10th, Doctorate, 5th-6th, Preschool.education-num: continuous
  "education.num", # continuous
  "marital.status", # Married-civ-spouse, Divorced, Never-married, Separated, Widowed, Married-spouse-absent, Married-AF-spouse.
  "occupation",  # Tech-support, Craft-repair, Other-service, Sales, Exec-managerial, Prof-specialty, Handlers-cleaners, Machine-op-inspct,                 # Adm-clerical, Farming-fishing, Transport-moving, Priv-house-serv, Protective-serv, Armed-Forces
  "relationship", # Wife, Own-child, Husband, Not-in-family, Other-relative, Unmarried
  "race", # White, Asian-Pac-Islander, Amer-Indian-Eskimo, Other, Black
  "sex",   # Female, Male
  "capital.gain", # continuous
  "capital.loss", # continuous
  "hours.per.week", # continuous
  "native.country", # United-States, Cambodia, England, Puerto-Rico, Canada, Germany, Outlying-US(Guam-USVI-etc), India, Japan, Greece, South, China, Cuba, Iran, Honduras, Philippines, Italy, Poland, Jamaica, Vietnam, Mexico, Portugal, Ireland, France, Dominican-Republic, Laos, Ecuador, Taiwan, Haiti, Columbia, Hungary, Guatemala, Nicaragua, Scotland, Thailand, Yugoslavia, El-Salvador, Trinadad&Tobago, Peru, Hong, Holand-Netherlands
  "income" # expects to predict
)
## First, convert "?"s to Not-Known...
adult.test[adult.test == "?"] <- "Not-Known"

# Similar decision is made to delete workclass that has Not-Known to utilize both workclass and occupation
adult.test = adult.test %>% filter(workclass!="Not-Known"& occupation!="Not-Known")

adult.test$education <- factor(adult.test$education, levels=edu.levels)
variables.to.factor = c("workclass","marital.status","occupation","relationship","race","sex","native.country","income")
adult.test[variables.to.factor] = lapply(adult.test[variables.to.factor],factor)

summary(adult.test)

#adult.missing = adult %>% filter_all(any_vars(is.na(.)))
#adult.test.missing = adult.test %>% filter_all(any_vars(is.na(.)))
# Look at missing data
library(naniar)
gg_miss_var(adult)
gg_miss_var(adult.test)

```

#### Before keeping the train/test split it came with, check out if training set has covered all the levels for test set
Category workclass, education, marital.status, occupation, relationship, race, native.country all need to be checked

**By Observing the given split below, all levels have been careful split to make sure all test set levels has been included in training set, and it looks like the split ratio is about 2:1 between train and test set. Decided to keep such split and proceed EDA on training set to prevent bias towards test set**

**Since the NA values only exist on occupation, workclass, and native.country, it make sense to keep them as Not Known instead of simply deleting them since there might be some common features among the Not Known categories to make the prediction more accurate - if they turn out to be outliers instead, we can always (went back to previous code chunk to rerun it to "Not Known" instead.)**

```{r}
summary(adult$workclass)
summary(adult.test$workclass)
# By Looking at workclass - it doesn't seem to have any level issues

summary(adult$education)
summary(adult.test$education)
# By Looking at education column - it doesn't seem to have any level issues

summary(adult$marital.status)
summary(adult.test$marital.status)
# By Looking at marital.status column - it doesn't seem to have any level issues

summary(adult$occupation)
summary(adult.test$occupation)
# By Looking at occupation column - it doesn't seem to have any level issues

summary(adult$relationship)
summary(adult.test$relationship)
# By Looking at relationship column - it doesn't seem to have any level issues

summary(adult$race)
summary(adult.test$race)
# By Looking at race column - it doesn't seem to have any level issues

summary(adult$native.country)
summary(adult.test$native.country)
# By Looking at native.country column - it doesn't seem to have any level issues
```

#### EDA on Training dataset - adult (initial inspection and correlation between predictors)
1. marital.status and relationship are correlated with each other - By looking at table marital.status vs relationship - their interaction can be potentially interesting in explaining a specific situation - might be interesting to explore as an interaction term
2. education.num and education are perfectly correlated with each other - education.num is probably directly derived from education level
3. Due to the fact missing values between workclass and occupation are almost identical and they seem to both contribute to predicting response variable (income), so decision is made to delete missing values in these two observations to preserve the ability to use both variables.
4. After identifying the correlation between the response above, by fitting a lm model for the remaining predictors, GVIF is no longer an issue

```{r}
library(GGally)
# education, occupation, and native-country have 15 levels or more - these levels need to be explored individually
# ggpairs(adult,columns=c(1,2,3,5,6,8,9,10,11,12,13),aes(colour=income))
# Comment out due to large run time. See image folder pairsplot1.png
# Blue is >50k, Orange <=50k
# Some workclass, marital.status, relationship, race, sex might be helpful for predicting income based on initial pairs plot

#create temp variable in order to fit lm model to test vif
temp = as.numeric(adult$income)
linear.model = lm(temp~.-income-education.num-occupation,data=adult)

library(car) # where vif lives
summary(linear.model)
vif(linear.model)

# See above result. Since occupation and education.num are creating aliased coefficients in the model they were removed from the  model first
# By looking at the rest of the predictors marital.status has GVIF of 63.54, and relationship has GVIF of 79.49 - these can potentiall have multicollinearity problems

# remove relationship variable see if GVIF gets better
linear.model1 = lm(temp~.-income-education.num-occupation-relationship,data=adult)
vif(linear.model1)

table(marital.status,relationship)
# Although by looking at the table above. It seems Marital status can be helpful in explaining status of a specific relationship status
# So even GVIF is high with them both present, their interaction can be helpful to make a better model

# looks like after relationship is removed marital.status becomes a lot better
# so marital.status and relationship are correlated with each other which also make intuitive sense

# Which variable is more associated with the response though? - try chi-squared test approximation(since dataset is quite large)
attach(adult)
chi.test <- chisq.test(table(income,marital.status))
chi.test
# Extremely small p-value < 2.2e-16 - marital.status and income are strongly associated

chi.test1 <- chisq.test(table(income,relationship))
chi.test1
# About the same levels of association
summary(marital.status)
summary(relationship)
# Can potentially explore either marital.status and relationship to see which one produce better prediction

# Note education.num and education may also be strongly correlated
plot(education.num~education)
# As we can see education.num and education are perfectly correlated with each other - education.num is probably directly derived from education level
# Although the actual education category doesn't assume equal distance between levels - it might be a better predictor overall

# Try getting rid of relationship and education.num and see if occupation is still an issue with GVIF
linear.model2 = lm(temp~.-income-education.num-relationship,data=adult)
summary(linear.model2)
vif(linear.model2)

# summary(linear.model2) tells us there might be a perfect correlation between Not-Known between workclass and occupation
# Following code to verify if this is true
adult.temp = adult %>% filter(workclass=="Not-Known"|occupation=="Not-Known") %>% select(workclass,occupation)
#view(adult.temp)
# There is a perfect correlation existed between missingi values of workclass and occupation in original dataset
summary(workclass)
summary(occupation)

# Take a look at whether workclass and occupation are independent from each other
chi.test2 <- chisq.test(table(occupation,workclass))
chi.test2
# So the two groups aren't independent from each other

# Check whether occupation contribute to predicting response variable
chi.test3 <- chisq.test(table(income,occupation))
chi.test3

# Check whether workclass contribute to predicting response variable
chi.test4 <- chisq.test(table(income,workclass))
chi.test4

# Both occupation and workclass seem to associated strongly with the response(income)
# It is best to go back to restrict our range of occupation and workclass to known only - since we want to use both to predict later on

# Looking at some categoricals predictors against response
table(income,sex) # sex can be contributing to income
table(income,workclass)
table(income,education) # Master, Prof-school, Doctorate all have more people making over 50k
table(income,marital.status)
table(income,relationship)
table(income,occupation)
table(income,race)
table(income,native.country)

```

#### EDA on Training dataset - adult (continue)
1. Based on the Pairs plot between principle components, there seem to have some seperation between two response levels under PC1 and PC2, PC1 and PC3, PC2 and PC3 seem to have more like a QDA relationship(QDA might run better)
2. Although there seem to be some seperation, this seperation doesn't seem to be particularly satisfying - so using continuous variables along may not lead to great prediction

```{r}
# It will be interesting to first perform a PCA and look at how numeric variables can contribute to predicting the response
reduced<-adult[,-c(2,4,6,7,8,9,10,14,15)]
# There are only five numeric variables including education.num which derived from education
pc.result<-prcomp(reduced,scale.=TRUE)
pc.scores<-pc.result$x

#Adding the response column to the PC's data frame
pc.scores<-data.frame(pc.scores)
pc.scores$income<-adult$income

#Use ggplot2 to plot the first few pc's
library(ggplot2)
ggplot(data = pc.scores, aes(x = PC1, y = PC2)) +
  geom_point(aes(col=income), size=1)+
  ggtitle("PCA of Census dataset")

ggplot(data = pc.scores, aes(x = PC2, y = PC3)) +
  geom_point(aes(col=income), size=1)+
  ggtitle("PCA of Census dataset")

# use ggpairs to plot all 6 components
# ggpairs(pc.scores,columns=c(1:6),aes(colour=income)) # Comment out due to slow plotting speed - image saved as PCA_pair1.png
# pairs plot between principle components - anaysis above this code chunk

# Explore some continuous variables predicting response
plot(age~income,col=c("red","blue")) # Might be a good predictor - mean are slighly different
plot(fnlwgt~income,col=c("red","blue")) # Almost identical between two groups - may not be a good predictor
plot(education.num~income,col=c("red","blue")) # education.num derived from education may be a good predictor if only continuous variable can be used
plot(capital.gain~income,col=c("red","blue")) # There is a floor value of 0 for both groups
summary(capital.gain)
have.capital.gain = adult %>% filter(capital.gain>0)
summary(have.capital.gain$capital.gain)
# view(have.capital.gain%>%select(income,capital.gain))
plot(have.capital.gain$capital.gain~have.capital.gain$income,col=c("red","blue")) 
# among the observations that have capital gain - capital gains is a good predictor
# seems capital.gain > 1st Qu, it is mostly over 50k income group
summary(have.capital.gain%>% filter(income==">50K")%>%select(capital.gain)) #1st Qu is 7289 capital gain - most people making over this threadhold is over 50k
capital.gain.over7298 = have.capital.gain%>%filter(capital.gain>=7298)
dim(capital.gain.over7298 %>% filter(income==">50K"))
dim(capital.gain.over7298 %>% filter(income=="<=50K"))
# Based on the result above - capital.gain over threadshold 7298 has 98.66% of people making over 50k
# This inspires us to create a indicator variable to improve accuracy rate

plot(capital.loss~income,col=c("red","blue")) #similar to capital gain most values are around 0
# However values that is not 0 seem to converge to similar mean - so it doesn't looks as helpful to us as a predictor
have.capital.loss = adult%>%filter(capital.loss>0)
plot(have.capital.loss$capital.loss~have.capital.loss$income,col=c("red","blue")) # mean seem to significantly different from each other

plot(hours.per.week~income,col=c("red","blue")) # hours.per.week can potentially be a good predictor since there seem to have a good seperation between mean of the two groups

# Do a pairwise plot to see if there are some multicollinearity between continuous variables
#ggpairs(reduced,aes(colour=income)) # commented out due to slow run time - under images continous_pair.png
```

### 2021-03-23 MSDS6372Project2.Rmd
wrap up EDA and finish Objective 1 coding

@@
 adult[variables.to.factor] = lapply(adult[variables.to.factor],factor)
+# Create variable cp.eover7298 per EDA discovery - capital gain equal or over 7298 - using "No" as reference
+adult$cp.eover7298<-factor(ifelse(adult$capital.gain>=7298,"Yes","No"),levels=c("No","Yes"))
 summary(adult)
@@
 adult.test[variables.to.factor] = lapply(adult.test[variables.to.factor],factor)
-
+# Create variable cp.eover7298 per EDA discovery - capital gain equal or over 7298 - using "No" as reference
+adult.test$cp.eover7298<-factor(ifelse(adult.test$capital.gain>=7298,"Yes","No"),levels=c("No","Yes"))
 summary(adult.test)
@@
 # Looking at some categoricals predictors against response
-table(income,sex) # sex can be contributing to income
-table(income,workclass)
-table(income,education) # Master, Prof-school, Doctorate all have more people making over 50k
-table(income,marital.status)
-table(income,relationship)
-table(income,occupation)
-table(income,race)
-table(income,native.country)
+
+ftable(addmargins(table(income,sex))) # sex can be contributing to income
+plot(income~sex,col=c("red","blue"))
+ftable(addmargins(table(income,workclass)))
+plot(income~workclass,col=c("red","blue"))
+ftable(addmargins(table(income,education))) # Master, Prof-school, Doctorate all have more people making over 50k
+plot(income~education,col=c("red","blue"))
+ftable(addmargins(table(income,marital.status)))
+plot(income~marital.status,col=c("red","blue"))
+ftable(addmargins(table(income,relationship)))
+plot(income~relationship,col=c("red","blue"))
+addmargins(table(income,occupation))
+plot(income~occupation,col=c("red","blue"))
+ftable(addmargins(table(income,race)))
+plot(income~race,col=c("red","blue"))
+addmargins(table(income,native.country))
+plot(income~native.country,col=c("red","blue"))
 
@@
 2. Although there seem to be some seperation, this seperation doesn't seem to be particularly satisfying - so using continuous variables along may not lead to great prediction
+3. Age and marrital status may have some interaction see below.
 
@@
 #ggpairs(reduced,aes(colour=income)) # commented out due to slow run time - under images continous_pair.png
+
+# Do some exploration between continuous variable and categorical predictors
+plot(age~education) # Not particularly interesting
+plot(hours.per.week~education) # Not particularly interesting
+plot(capital.gain~cp.eover7298) # variable created 
+plot(age~marital.status) # Widowed have the highest mean age
+plot(age~relationship)
+
 ```
+
+#### Objective 1 - Build a simple Logistic regression model for interpretation
+
+```{r}
+library(MASS)
+library(glmnet)
+dat.adult.x <- model.matrix(income~.-1,adult)
+dat.adult.y<-adult[,15]
+cvfit <- cv.glmnet(dat.adult.x, dat.adult.y, family = "binomial", type.measure = "class", nlambda = 1000)
+plot(cvfit)
+coef(cvfit, s = "lambda.min")
+print("Penalty Value:")
+cvfit$lambda.min
+# By Looking at the turned on coefficient at lambda.min:
+# age, workclass, education, education.num, marital.status, occupation, relationship, race, sex, capital.gain, capital.loss
+# hours.per.week, native.country, cp.eover7298
+# Are all turned on
+##############################################
+# However we know from our EDA
+# education and education.num are perfectly correlated with each other - so we will keep education only(delete education.num from simple model)
+# cp.eover7298 is more useful variable that derived from capital.gain - so we will keep cp.eover7298 instead of capital.gain
+# marital status has four levels turned on, while relationship also has 4 levels turned on - however marital status has bigger coefficient(so we will keep marital status in the model and get rid of relationship variable for now since we don't want GVIF issue with a interpretable model)
+
+# Following refit with glm model with the variables turned on - EDA variables we deemed less important
+simple.log <- glm(income~age+workclass+education+marital.status+occupation+race+sex+capital.loss+hours.per.week+native.country+cp.eover7298,family="binomial",data=adult)
+step.log<-simple.log %>% stepAIC(trace=FALSE)
+summary(step.log)
+exp(cbind("Odds ratio" = coef(step.log), confint.default(step.log, level = 0.95)))
+
+# As we can see from above stepAIC function got rid of native.country variable - it has too many levels and the training model seem to find better balance without it
+# Even education is not significantly different from reference level of Preschool - the group as a whole has been kept by stepAIC
+# Furthermore - education's coefficient, eg: educationDoctorate = 15.5 - seem to contribute a lot more than say workclassState-gov -0.84
+# So we will keep education in the model and refit the model
+
+simple.log1 <- glm(income~age+workclass+education+marital.status+occupation+race+sex+capital.loss+hours.per.week+cp.eover7298,family="binomial",data=adult)
+summary(simple.log1)
+exp(cbind("Odds ratio" = coef(simple.log1), confint.default(simple.log1, level = 0.95)))
+
+# Look at the leverage and cook's D plot for the model
+plot(simple.log1)
+# As we can see from the Residuals vs Leverage plot - there are some higher leverage points
+# However Cook's distance suggests no observation seem to be of great concern in our model(no high residual and high leverage point)
+
+#Making predictions for simple.log1 for ROC plot
+fit.pred.sm1<-predict(simple.log1,newdata=adult.test,type="response")
+
+# Generate ROC curve to find a cutoff to use
+library(ROCR)
+results.sm1 <-prediction(fit.pred.sm1, adult.test$income,label.ordering=c("<=50K.",">50K."))
+roc.sm1 = performance(results.sm1, measure = "tpr", x.measure = "fpr")
+plot(roc.sm1,colorize = TRUE)
+abline(a=0, b= 1)
+
+# Since this dataset is to predict whether income is less than or equal to 50K or greater than 50K
+# We want the sensitivity and specificity to be roughly the same - no reasoon to prefer one group or the other
+# As we can see from the color legend, the greant to light blue region will give over 80% True positive rate and less than 20% False positive rate
+# And that cut off is roughly correspond to 0.275
+
+#Calculate logistic regression using cut of 0.3 on test set
+cutoff<-0.275
+class.sm1<-factor(ifelse(fit.pred.sm1>cutoff,">50K.","<=50K."),levels=c("<=50K.",">50K."))
+library(caret)
+confusionMatrix(class.sm1,adult.test$income)
+
+# As we can see the Sensitivity and Specificity is roughly the same both over 80%
+
+```

### 2021-03-23 MSDS6372Project2.Rmd
add comment on continuous variable

@@
 3. Age and marrital status may have some interaction see below.
+4. We didn't find any highly correlated continous variables
 
@@
 # Do a pairwise plot to see if there are some multicollinearity between continuous variables
-#ggpairs(reduced,aes(colour=income)) # commented out due to slow run time - under images continous_pair.png
+#ggpairs(reduced,aes(colour=income)) # commented out due to slow run time - under images continuous_pair.png
+# We didn't find any highly correlated continous variables
 

### 2021-03-23 MSDS6372Project2.Rmd
Object2 part1 - try building complex model

@@
 2. Although there seem to be some seperation, this seperation doesn't seem to be particularly satisfying - so using continuous variables along may not lead to great prediction
-3. Age and marrital status may have some interaction see below.
+3. Age and marital.status may have some interaction see below.
 4. We didn't find any highly correlated continous variables
@@
 plot(age~relationship)
+plot(sex~education)
+plot(sex~occupation) # Sex and occupation might be interesting interaction term
 
@@
 # As we can see the Sensitivity and Specificity is roughly the same both over 80%
+# Accuracy : 0.8184 Sensitivity : 0.8226 Specificity : 0.8057
 
 ```
+
+#### Objective 2 Part 1 - more complicated logistic regression model
+```{r}
+# With our EDA age and marital status might be interesting to explore as an interaction
+# Also relationship and marital.status might also be useful to be considered as interaction
+# Run LASSO to see what variables turn on or off
+dat.adult.x1 <- model.matrix(income~age+workclass+education+marital.status+occupation+race+sex+capital.loss+hours.per.week+cp.eover7298+age:marital.status+relationship+relationship:marital.status-1,adult)
+dat.adult.y1<-adult[,15]
+cvfit1 <- cv.glmnet(dat.adult.x1, dat.adult.y1, family = "binomial", type.measure = "class", nlambda = 1000)
+plot(cvfit1)
+coef(cvfit1, s = "lambda.min")
+print("Penalty Value:")
+cvfit1$lambda.min
+# age:marital.status didn't get turned on
+# some levels under marital.status interact with relationship got turned on
+
+# We will also add native.country back in to see if it can improve performance 
+# Fit a glm model by taking out age:marital.status but keep marital.status:relationship
+# Run a forward selection to see if anything get kept in the model
+complex.log<-glm(income~age+workclass+education+marital.status+occupation+race+sex+capital.loss+hours.per.week+cp.eover7298+relationship+relationship:marital.status+native.country,family="binomial",data=adult)
+step.log1<-complex.log %>% stepAIC(trace=FALSE)
+summary(step.log1)
+
+# We can see from the stepAIC selection
+# age, workclass, education, marital.status, occupation, race, sex, capital.loss, hours.per.week, cp.eover7298, relationship, native.country got selected
+# Let's compare both complex.log and step.log1 prediction performance on the test set
+
+#Making predictions for complex.log for ROC plot#####
+fit.pred.complex1<-predict(complex.log,newdata=adult.test,type="response")
+
+# Generate ROC curve to find a cutoff to use
+results.complex1 <-prediction(fit.pred.complex1, adult.test$income,label.ordering=c("<=50K.",">50K."))
+roc.complex1 = performance(results.complex1, measure = "tpr", x.measure = "fpr")
+plot(roc.complex1,colorize = TRUE)
+abline(a=0, b= 1)
+#Calculate logistic regression complex using cut of 0.3 on test set
+cutoff1<-0.275
+class.complex1 <-factor(ifelse(fit.pred.complex1>cutoff1,">50K.","<=50K."),levels=c("<=50K.",">50K."))
+confusionMatrix(class.complex1,adult.test$income)
+# The complex model has about the same performance by adding in relationship:marital.status interaction
+# Accuracy : 0.817  Sensitivity : 0.8192 Specificity : 0.8102
+
+#Making predictions for step.log1 for ROC plot#####
+
+fit.pred.step1<-predict(step.log1,newdata=adult.test,type="response")
+
+# Generate ROC curve to find a cutoff to use
+results.step1 <-prediction(fit.pred.step1, adult.test$income,label.ordering=c("<=50K.",">50K."))
+roc.step1 = performance(results.step1, measure = "tpr", x.measure = "fpr")
+plot(roc.step1,colorize = TRUE)
+abline(a=0, b= 1)
+#Calculate logistic regression complex using cut of 0.3 on test set
+cutoff1<-0.275
+class.step1 <-factor(ifelse(fit.pred.step1>cutoff1,">50K.","<=50K."),levels=c("<=50K.",">50K."))
+confusionMatrix(class.step1,adult.test$income)
+# The step1 model has about the same performance compared to complex model with interaction
+# Accuracy : 0.8174 Sensitivity : 0.8194 Specificity : 0.8112
+
+# Now exploring adding sex and occupation interaction
+complex.log2<-glm(income~age+workclass+education+marital.status+occupation+race+sex+capital.loss+hours.per.week+cp.eover7298+relationship+native.country+sex:occupation,family="binomial",data=adult)
+
+#Making predictions for complex.log2 for ROC plot#####
+fit.pred.complex2<-predict(complex.log2,newdata=adult.test,type="response")
+
+# Generate ROC curve to find a cutoff to use
+results.complex2 <-prediction(fit.pred.complex2, adult.test$income,label.ordering=c("<=50K.",">50K."))
+roc.complex2 = performance(results.complex2, measure = "tpr", x.measure = "fpr")
+plot(roc.complex2,colorize = TRUE)
+abline(a=0, b= 1)
+#Calculate logistic regression complex using cut of 0.3 on test set
+cutoff1<-0.275
+class.complex2 <-factor(ifelse(fit.pred.complex2>cutoff1,">50K.","<=50K."),levels=c("<=50K.",">50K."))
+confusionMatrix(class.complex2,adult.test$income)
+# The complex model has about the same performance by adding in relationship:marital.status interaction
+# Accuracy : Accuracy : 0.8167  Sensitivity : 0.8183 Specificity : 0.8118
+
+```

### 2021-03-23 MSDS6372Project2.Rmd
objective 2 and beyond

@@
 abline(a=0, b= 1)
-
+auc.sm1 <- performance(results.sm1, measure = "auc")
+auc.sm1 <- auc.sm1@y.values
+text(x = .40, y = .6,paste("AUC = ", round(auc.sm1[[1]],3), sep = ""))
+# AUC = 0.903
 # Since this dataset is to predict whether income is less than or equal to 50K or greater than 50K
@@
 abline(a=0, b= 1)
+auc.complex1 <- performance(results.complex1, measure = "auc")
+auc.complex1 <- auc.complex1@y.values
+text(x = .40, y = .6,paste("AUC = ", round(auc.complex1[[1]],3), sep = ""))
+# AUC = 0.903
 #Calculate logistic regression complex using cut of 0.3 on test set
@@
 abline(a=0, b= 1)
+auc.step1 <- performance(results.step1, measure = "auc")
+auc.step1 <- auc.step1@y.values
+text(x = .40, y = .6,paste("AUC = ", round(auc.step1[[1]],3), sep = ""))
+# AUC = 0.904
 #Calculate logistic regression complex using cut of 0.3 on test set
@@
 abline(a=0, b= 1)
-#Calculate logistic regression complex using cut of 0.3 on test set
+auc.complex2 <- performance(results.complex2, measure = "auc")
+auc.complex2 <- auc.complex2@y.values
+text(x = .40, y = .6,paste("AUC = ", round(auc.complex2[[1]],3), sep = ""))
+# AUC = 0.904
+#Calculate logistic regression complex using cut of 0.275 on test set
 cutoff1<-0.275
@@
 ```
+
+#### Objective 2 Part 2 - Create competing model using continuous predictors and use LDA or QDA
+LDA model is the winner after comparison:
+Even QDA have slightly better AUC value(suspect QDA is overfit) - the cut off of probability is extremely sensitive(highly dependent on test set)
+Thus LDA is the winner
+However, LDA performance isn't as good compared to Logistic Regression
+
+```{r}
+# construct the LDA model
+# Fron EDA we know fnlwgt is not helpful in identifying income - will not use it here
+# We will use education.num since it is a substitute for education
+mylda <- lda(income ~ age + education.num+ capital.gain+capital.loss+hours.per.week, data = adult)
+mylda.prd <- predict(mylda, newdata = adult.test)
+mylda.posterior <- mylda.prd$posterior[,2]
+
+# Generate ROC curve to find a cutoff to use
+results.lda <-prediction(mylda.posterior, adult.test$income,label.ordering=c("<=50K.",">50K."))
+roc.lda = performance(results.lda, measure = "tpr", x.measure = "fpr")
+plot(roc.lda,colorize = TRUE)
+abline(a=0, b= 1)
+auc.lda <- performance(results.lda, measure = "auc")
+auc.lda <- auc.lda@y.values
+text(x = .40, y = .6,paste("AUC = ", round(auc.lda[[1]],3), sep = ""))
+# As we can see the lda model is slighly less good compared to logistic regression model
+# The ROC curve has shifted downwards compared to LR - AUC =0.809(decreased)
+# Nevertheless to balance off performance between high True positive and low False positive we can use around 0.25
+
+#Calculate logistic regression complex using cut of 0.25 on test set
+cutoff1<-0.23
+class.lda <-factor(ifelse(mylda.posterior>cutoff1,">50K.","<=50K."),levels=c("<=50K.",">50K."))
+confusionMatrix(class.lda,adult.test$income)
+# Accuracy : Accuracy : 0.726  Sensitivity : 0.7263 Specificity : 0.7251
+
+##################################################################################################
+# construct the QDA model
+# Fron EDA we know fnlwgt is not helpful in identifying income - will not use it here
+# We will use education.num since it is a substitute for education
+myqda <- qda(income ~ age + education.num+ capital.gain+capital.loss+hours.per.week, data = adult)
+myqda.prd <- predict(myqda, newdata = adult.test)
+myqda.posterior <- myqda.prd$posterior[,2]
+
+# Generate ROC curve to find a cutoff to use
+results.qda <-prediction(myqda.posterior, adult.test$income,label.ordering=c("<=50K.",">50K."))
+roc.qda = performance(results.qda, measure = "tpr", x.measure = "fpr")
+plot(roc.qda,colorize = TRUE)
+abline(a=0, b= 1)
+auc.qda <- performance(results.qda, measure = "auc")
+auc.qda <- auc.qda@y.values
+text(x = .40, y = .6,paste("AUC = ", round(auc.qda[[1]],3), sep = ""))
+# AUC seem to be slightly better than LDA AUC = 0.823
+# To balance off performance between high True positive and low False positive we can use around 0.25
+
+#Calculate logistic regression complex using cut of 0.015 on test set
+cutoff2<-0.015
+class.qda <-factor(ifelse(myqda.posterior>cutoff2,">50K.","<=50K."),levels=c("<=50K.",">50K."))
+confusionMatrix(class.qda,adult.test$income)
+# Accuracy : Accuracy : 0.7461  Sensitivity : 0.7563 Specificity : 0.7150
+# Due to the extreme cut off we know even QDA has better AUC value, the prediction cut off is extremely sensitive based on the test set
+# or future validation set
+# LDA model is much more stable even the AUC value is slightly less
+
+```
+
+#### Objective 2 part 3 - Nonparametric model approach - Random forest
+Default RF ntree = 10 and m variables =sqrt(p) seem to work better than ntree=100
+```{r}
+library(randomForest)
+set.seed(1)
+# By default use sqrt(p) number variables when building a random forest of classification trees
+rf.adult = randomForest(income~.,data=adult)
+
+# Fix variable levels in adult.test in order to use randomForest
+levels(adult.test$workclass)<-levels(adult$workclass)
+levels(adult.test$education)<-levels(adult$education)
+levels(adult.test$marital.status)<-levels(adult$marital.status)
+levels(adult.test$occupation)<-levels(adult$occupation)
+levels(adult.test$relationship)<-levels(adult$relationship)
+levels(adult.test$race)<-levels(adult$race)
+levels(adult.test$native.country)<-levels(adult$native.country)
+
+importance(rf.adult)
+plot(rf.adult)
+# Run randomForest prediction
+testdata = adult.test[,-15]
+rf.pred.prob = predict(rf.adult,newdata=testdata,type = "prob")
+rf.prob = rf.pred.prob[,2]
+
+# Generate ROC curve to find a cutoff to use
+results.rf <-prediction(rf.prob, adult.test$income,label.ordering=c("<=50K.",">50K."))
+roc.rf = performance(results.rf, measure = "tpr", x.measure = "fpr")
+plot(roc.rf,colorize = TRUE)
+abline(a=0, b= 1)
+auc.rf <- performance(results.rf, measure = "auc")
+auc.rf <- auc.rf@y.values
+text(x = .40, y = .6,paste("AUC = ", round(auc.rf[[1]],3), sep = ""))
+# AUC seem to be the best yet AUC = 0.907
+# To balance off performance between high True positive and low False positive we can use around 0.25
+
+#Calculate logistic regression complex using cut of 0.425 on test set
+cutoff3<-0.425
+class.rf <-factor(ifelse(rf.prob>cutoff3,">50K.","<=50K."),levels=c("<=50K.",">50K."))
+confusionMatrix(class.rf,adult.test$income)
+# Accuracy : 0.8114 Sensitivity : 0.8118 Specificity : 0.8102
+
+###############################################################################################
+
+set.seed(1)
+# By default use sqrt(p) number variables when building a random forest of classification trees
+# Try Number of Trees to 100
+rf.adult1 = randomForest(income~.,data=adult,ntree=100)
+
+importance(rf.adult1)
+plot(rf.adult1)
+# Run randomForest prediction
+rf.pred.prob1 = predict(rf.adult1,newdata=testdata,type = "prob")
+rf.prob1 = rf.pred.prob1[,2]
+
+# Generate ROC curve to find a cutoff to use
+results.rf1 <-prediction(rf.prob1, adult.test$income,label.ordering=c("<=50K.",">50K."))
+roc.rf1 = performance(results.rf1, measure = "tpr", x.measure = "fpr")
+plot(roc.rf1,colorize = TRUE)
+abline(a=0, b= 1)
+auc.rf1 <- performance(results.rf1, measure = "auc")
+auc.rf1 <- auc.rf1@y.values
+text(x = .40, y = .6,paste("AUC = ", round(auc.rf1[[1]],3), sep = ""))
+# AUC seem to be the best yet AUC = 0.902
+# To balance off performance between high True positive and low False positive we can use around 0.25
+
+#Calculate logistic regression complex using cut of 0.43 on test set
+cutoff4<-0.41
+class.rf1 <-factor(ifelse(rf.prob1>cutoff4,">50K.","<=50K."),levels=c("<=50K.",">50K."))
+confusionMatrix(class.rf1,adult.test$income)
+# Accuracy : 0.8089 Sensitivity : 0.8114 Specificity : 0.8012
+
+```
+
+#### Overlay multiple ROC curve together
+```{r}
+plot(roc.sm1)
+plot(roc.complex2,col="orange", add = TRUE)
+plot(roc.lda,col="blue",add=TRUE)
+plot(roc.rf,col="green",add=TRUE)
+legend("bottomright",legend=c("Simple LR","Complex LG","LDA","Random Forest"),col=c("black","orange","blue","green"),lty=1,lwd=1)
+abline(a=0, b= 1)
+
+```

### 2021-03-23 MSDS6372Project2.Rmd
minor adjustment

@@
 plot(roc.rf,col="green",add=TRUE)
-legend("bottomright",legend=c("Simple LR","Complex LG","LDA","Random Forest"),col=c("black","orange","blue","green"),lty=1,lwd=1)
+legend("bottomright",legend=c("Simple LR","Complex LR","LDA","Random Forest"),col=c("black","orange","blue","green"),lty=1,lwd=1)
 abline(a=0, b= 1)

### 2021-03-27 MSDS6372Project2.Rmd
minoro update

@@
 
+######## This Code part has been readdressed at the above importing section #############################################
 # summary(linear.model2) tells us there might be a perfect correlation between Not-Known between workclass and occupation
 # Following code to verify if this is true
-adult.temp = adult %>% filter(workclass=="Not-Known"|occupation=="Not-Known") %>% select(workclass,occupation)
+#adult.temp = adult %>% filter(workclass=="Not-Known"|occupation=="Not-Known") %>% select(workclass,occupation)
 #view(adult.temp)
+########################################################################################################################
 # There is a perfect correlation existed between missingi values of workclass and occupation in original dataset
@@
 # It will be interesting to first perform a PCA and look at how numeric variables can contribute to predicting the response
-reduced<-adult[,-c(2,4,6,7,8,9,10,14,15)]
+reduced<-adult[,-c(2,4,6,7,8,9,10,14,15,16)]
 # There are only five numeric variables including education.num which derived from education

### 2021-03-29 MSDS6372Project2.Rmd
update to objective 1

@@
 summary(simple.log1)
+vif(simple.log1)
 exp(cbind("Odds ratio" = coef(simple.log1), confint.default(simple.log1, level = 0.95)))
@@
 
+# Try another model without education since it is not significant
+simple.log2 <- glm(income~age+workclass+marital.status+occupation+race+sex+capital.loss+hours.per.week+cp.eover7298,family="binomial",data=adult)
+summary(simple.log2)
+plot(simple.log2)
+fit.pred.sm2<-predict(simple.log2,newdata=adult.test,type="response")
+
+results.sm2 <-prediction(fit.pred.sm2, adult.test$income,label.ordering=c("<=50K.",">50K."))
+roc.sm2 = performance(results.sm2, measure = "tpr", x.measure = "fpr")
+plot(roc.sm2,colorize = TRUE)
+abline(a=0, b= 1)
+auc.sm2 <- performance(results.sm2, measure = "auc")
+auc.sm2 <- auc.sm2@y.values
+text(x = .40, y = .6,paste("AUC = ", round(auc.sm2[[1]],3), sep = ""))
+# AUC = 0.892
+# Since this dataset is to predict whether income is less than or equal to 50K or greater than 50K
+# We want the sensitivity and specificity to be roughly the same - no reasoon to prefer one group or the other
+# As we can see from the color legend, the greant to light blue region will give over 80% True positive rate and less than 20% False positive rate
+# And that cut off is roughly correspond to 0.275
+
+#Calculate logistic regression using cut of 0.3 on test set
+cutoff<-0.26
+class.sm2<-factor(ifelse(fit.pred.sm2>cutoff,">50K.","<=50K."),levels=c("<=50K.",">50K."))
+confusionMatrix(class.sm2,adult.test$income)
+vif(simple.log2)
+#Accuracy : 0.8048 Sensitivity : 0.8058 Specificity : 0.8020
 ```

### 2021-04-06 MSDS6372Project2.Rmd
update random forest with different trial

@@
 #### Objective 2 part 3 - Nonparametric model approach - Random forest
-Default RF ntree = 10 and m variables =sqrt(p) seem to work better than ntree=100
+- Default RF ntree = 500 and m variables =sqrt(p) seem to work better than ntree=100
+- Try iterate through mtry variable from 1 to 10 using prediction cut off previously discovered - seems mtry = 1 works best on overall accuracy
+- After actually fitting random forest with mtry = 1 the AUC = 0.882 (performance is worse compare to default mtry=3).
+Sensitivity and Specificity also suffers after adjusting the cut off.
+So Default RF option still wins!
+- So Mtry = 3 seems to be the perfect balance for our dataset in order to provide good AUC.
 ```{r}
@@
 class.rf <-factor(ifelse(rf.prob>cutoff3,">50K.","<=50K."),levels=c("<=50K.",">50K."))
-confusionMatrix(class.rf,adult.test$income)
+cm = confusionMatrix(class.rf,adult.test$income)
 # Accuracy : 0.8114 Sensitivity : 0.8118 Specificity : 0.8102
@@
 
+##### Find best mtry ####
+# Use custom loop to find it based on training model
+# Using previous discovered cutoff3 = 0.425 
+parameter = 10
+overall.Accuracy = c()
+for(i in 1:parameter){
+  set.seed(1)
+  rf.adult.temp = randomForest(income~.,data=adult,mtry=i)
+  rf.pred.temp = predict(rf.adult.temp,newdata=testdata,type = "prob")
+  rf.prob.temp = rf.pred.temp[,2]
+  class.rf.temp <-factor(ifelse(rf.prob.temp>cutoff3,">50K.","<=50K."),levels=c("<=50K.",">50K."))
+  cm.temp = confusionMatrix(class.rf.temp,adult.test$income)
+  overall.Accuracy[i]=cm.temp$overall[1]
+}
+overall.Accuracy
+
+###### Giving us Mtry = 1 is the best split ####
+# Try this option
+
+set.seed(1)
+rf.adult2 = randomForest(income~.,data=adult,mtry=1)
+
+# Run randomForest prediction
+rf.pred.prob2 = predict(rf.adult2,newdata=testdata,type = "prob")
+rf.prob2 = rf.pred.prob2[,2]
+
+# Generate ROC curve to find a cutoff to use
+results.rf2 <-prediction(rf.prob2, adult.test$income,label.ordering=c("<=50K.",">50K."))
+roc.rf2 = performance(results.rf2, measure = "tpr", x.measure = "fpr")
+plot(roc.rf2,colorize = TRUE)
+abline(a=0, b= 1)
+auc.rf2 <- performance(results.rf2, measure = "auc")
+auc.rf2 <- auc.rf2@y.values
+text(x = .40, y = .6,paste("AUC = ", round(auc.rf2[[1]],3), sep = ""))
+# AUC seem to be the best yet AUC = 0.882
+# To balance off performance between high True positive and low False positive we can use around 0.2
+
+#Calculate logistic regression complex using cut of 0.43 on test set
+cutoff5<-0.16
+class.rf2 <-factor(ifelse(rf.prob2>cutoff5,">50K.","<=50K."),levels=c("<=50K.",">50K."))
+confusionMatrix(class.rf2,adult.test$income)
+
 ```

### 2021-04-06 MSDS6372Project2.Rmd
more update on mtry rf

@@
 - So Mtry = 3 seems to be the perfect balance for our dataset in order to provide good AUC.
+- We further added in calculations for AUC under each mtry #, and it also tells us mtry=3 provides the best AUC value
+> overall.Accuracy
+ [1] 0.8526934 0.8479922 0.8114267 0.7909892 0.7781913 0.7676787 0.7533790 0.7478289 0.7484166 0.7407770
+> overall.auc
+ [1] 0.882 0.903 0.907 0.903 0.899 0.893 0.890 0.887 0.883 0.878
 ```{r}
@@
 # Using previous discovered cutoff3 = 0.425 
+# Adjust this to also find best mtry for AUC
 parameter = 10
 overall.Accuracy = c()
+overall.auc = c()
 for(i in 1:parameter){
+  print(paste("Processing mtry=",i))
   set.seed(1)
@@
   overall.Accuracy[i]=cm.temp$overall[1]
+
+  # Calculate AUC
+  results.rf.temp <-prediction(rf.prob.temp, adult.test$income,label.ordering=c("<=50K.",">50K."))
+  auc.temp <- performance(results.rf.temp, measure = "auc")
+  auc.temp <- auc.temp@y.values
+  overall.auc[i] <- round(auc.temp[[1]],3)
 }
 overall.Accuracy
+overall.auc
 
-###### Giving us Mtry = 1 is the best split ####
+###### Giving us Mtry = 1 is the best split based on overall.Accuracy ####
 # Try this option

### 2021-04-10 MSDS6372Project2.Rmd
bug fix

@@
  [1] 0.882 0.903 0.907 0.903 0.899 0.893 0.890 0.887 0.883 0.878
+
 ```{r}
